### 2016-05-21 writing/latex/coincidence.tex
Add changes to final section
Slides and fixes t bibliography
Changes to presentation from Henning's notes

\begin{lemma}
    \label{lem:incdec}
    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
    in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
    \(\pi \preceq \adda{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
    then in any occurrence of \(m\) in a permutation \(\varrho\) the region
    corresponding to the box \((a,b)\) can only contain an increasing
    (decreasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

@@

\end{tikzpicture}
    \caption{If the conditions of \PropositionRef{prop:dom2} are satisfied the box
    \((a-1,b)\) can be shaded.}
\end{figure}

@@

Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
If the region is empty then we have an occurrence of \(m_2\).
If there is only one point in the region we can choose this to replace the \(1\) in
the mesh pattern to get the required shading. If there is more than one point then
choosing the two leftmost points gives us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
Now if we take the other two points as the points in our
mesh pattern then we get an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.

>>>>

\begin{lemma}
    \label{lem:incdec}
    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
    in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
    \(\pi \preceq \adda{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
    ,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
    corresponding to the box \((a,b)\) can only contain an decreasing
    (increasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

@@

\end{tikzpicture}
    \caption{If the conditions of \PropositionRef{prop:dom2} are satisfied the box
    \((a,b)\) can be shaded.}
\end{figure}

@@

Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
If the region is empty then we have an occurrence of \(m_2\).
If there is only one point in the region we can choose this to replace the \(1\) in
the mesh pattern to get the required shading. If there is more than one point then
choosing the two leftmost points gives us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
Now if we take the other two points as the points in our
mesh pattern then we get an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.
It is also possible to calculate this coincidence by an extension of the Third
Dominating rule, where we allow a sequence of \texttt{add\_point} operations, this
is discussed further in the future work section.

### 2016-05-21 writing/presentation/coincidences.tex
Changes to presentation from Henning's notes

\section{Coincidence Classes}
\label{sec:Coincidence Classes}
\subsection{Coincidence}
\label{sub:Coincidence}
\begin{frame}{Coincidence}
  \begin{definition}
    Two mesh patterns are said to be \emph{coincident} if they avoid the
    same set of permutations at every length.
  \end{definition}
  \begin{block}{}
    Classical patterns can never be coincident.

@@

\begin{frame}
  \begin{lemma}
    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
\(\pi \preceq \adda{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
corresponding to the box \((a,b)\) can only contain an increasing
(decreasing) subsequence of \(\varrho\).
  \end{lemma}
\end{frame}

@@

\subsection{Special Cases}
\label{sub:Special Cases}
\begin{frame}{Special Cases}
  There are \(7\) coincidences of the form \(\av{\perm{3,2,1},m}\) that are not
  explained by the rules.
  \begin{example}
    \begin{equation*}
      m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
    \end{equation*}
    This coincidence is explained by mathematical induction on the number of
    points in region corresponding to the middle box.
  We call this number \(n\).
  \end{example}
\end{frame}
\begin{frame}
  \begin{example}
    \begin{description}
    \only<1-2>{\item [Base Case \((n=0)\):] The base case holds since we can freely shade the
        box if it contains no points.}
    \only<2>{\item [Inductive Hypothesis \((n=k)\):] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.}
    \only<3>{\item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, this gives the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done.}
        \only<4>{\item [Inductive Step \((n=k+1)\) (cont.)] If this box contains
        any points then we gain some extra shading on the mesh pattern
        due to the dominating pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.}
\end{description}
  \end{example}
\end{frame}

>>>>

\section{Coincidence Classes}
\label{sec:Coincidence Classes}
\subsection{Coincidence}
\label{sub:Coincidence}
\begin{frame}{Coincidence}
  \begin{definition}
    Two mesh patterns are said to be \emph{coincident} if they avoid the
    same set of permutations at every length.
  \end{definition}
  \begin{block}{}
    Distinct classical patterns can never be coincident.

@@

\begin{frame}
  \begin{lemma}
    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
\(\pi \preceq \adda{\sigma}{(a,b)}\)\\
,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
corresponding to the box \((a,b)\) can only contain an decreasing
subsequence of \(\varrho\).
  \end{lemma}
\end{frame}

@@

\subsection{Special Cases}
\label{sub:Special Cases}
\begin{frame}{Special Cases}
  There are \(7\) coincidences of the form \(\av{\perm{3,2,1},m}\) that are not
  explained by the rules.
  \begin{example}
    \begin{equation*}
      m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
    \end{equation*}
    This coincidence is explained by mathematical induction on the number of
    points in the region corresponding to the middle box.
  We call this number \(n\).
  \end{example}
\end{frame}
\begin{frame}
  \begin{example}
    \begin{description}
    \only<1-2>{\item [Base Case \((n=0)\):] The base case holds since we can freely shade the
        box if it contains no points.}
    \only<2>{\item [Inductive Hypothesis \((n=k)\):] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.}
    \only<3>{\item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, this gives the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done.}
        \only<4>{\item [Inductive Step \((n=k+1)\) (cont.)] If this box contains
        any points then we gain some extra shading on the mesh pattern
        due to the dominating pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.}
\end{description}
  \end{example}
\end{frame}

### 2016-05-21 writing/presentation/introduction.tex
Changes to presentation from Henning's notes

  We write permutations in \emph{one-line notation}, writing the entries of
  the entries of the permutation in order
  \begin{equation*}
    \pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
  \end{equation*}
  \begin{example} The \(6\) permutations on \(\nrange{3}\) are
    \begin{equation*}
      123, 132, 213, 231, 312, 321
    \end{equation*}
  \end{example}
\end{frame}

@@

  We the set of permutations of length \(n\) avoiding a pattern \(\sigma\) is
  denoted as \(\Av_n(\sigma)\) and
  \(\av{\sigma} = \bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
\end{frame}

>>>>

  We write permutations in \emph{one-line notation}, writing
  the entries of the permutation in order
  \begin{equation*}
    \pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
  \end{equation*}
  \begin{example} The \(6\) permutations on \(\nrange{3}\) are
    \begin{equation*}
      123, 132, 213, 231, 312, 321
    \end{equation*}
  \end{example}
\end{frame}

@@

  The set of permutations of length \(n\) avoiding a pattern \(\sigma\) is
  denoted as \(\Av_n(\sigma)\) and
  \begin{equation*}
    \av{\sigma} = \bigcup_{i=0}^{\infty}\Av_i(\sigma)
  \end{equation*}
\end{frame}

### 2016-05-21 writing/presentation/wilf.tex
Changes to presentation from Henning's notes

\section{Wilf-equivalences}
\label{sec:Wilf-equivalences}
\subsection{Wilf-equivalence}
\label{sub:Wilf-equivalence}
\begin{frame}{Wilf-equivalence}
  \begin{definition}
  Two patterns are said to be \emph{Wilf-equivalent} if they avoid (or contain)
  the same number of patterns at every length.
\end{definition}
 All coincident pattern classes are Wilf-equivalent.

 The permutations in \(\av{\perm{2,3,1}}\) of length \(n\) are counted by the
 Catalan numbers.
 \begin{equation}
   C_0 = 1, \text{ and } C_{n+1} =\sum_{i=0}^{n}C_iC_{n-i}
 \end{equation}
 This gives the sequence
 \begin{equation*}
   1,1,2,5,14,42,132,1430,4862,\dotsc
 \end{equation*}
\end{frame}
\subsection{Set Partitions}
\label{sub:Set Partitions}
\begin{frame}{Set Partitions}
  \begin{block}{Note}
    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
    \(\nrange{n}\). (Claesson)
\end{block}
\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
\end{example}
All permutations in \(\av{\perm{2,3,1}}\) are also in \(\av{q}\).
\end{frame}
\begin{frame}
  Set Partitions can be used to explain certain Wilf-equivalences.
  \begin{example}
    The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
  \begin{equation*}
      m_1 = \pattern{}{1,2}{0/0,0/2,
                      1/0,1/1,
                      2/0,2/1,2/2}
      \text{ and }
      m_2 = \pattern{}{1,2}{0/0,
                      1/0,1/1,1/2,
                      2/0,2/1,2/2}
  \end{equation*}
  Consider containers of the patterns in \(\av{\perm{2,3,1}}\).

@@

\begin{frame}
  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\mathcal{M}_2}
\end{equation*}
Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\scriptstyle \mathcal{M}_2\setminus\varepsilon}
\end{equation*}
\end{frame}
\begin{frame}
  Both of these sets therefore have generating function \(M(x)\) satisfying
  \begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}\pause
 Solving Equation~\ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
Equation~\ref{eqn:Pgen} gives us that the generating function for
\(M(x)\) satisfies
\begin{equation}
    M(x) = xM^2(x) - x(M(x) - 1) + 1
\end{equation}\pause
Evaluating coefficients of \(M(x)\) then gives the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
\end{equation*}
This is an offset of the Motzkin numbers.
\end{frame}

>>>>

\section{Wilf-equivalences}
\label{sec:Wilf-equivalences}
\subsection{Wilf-equivalence}
\label{sub:Wilf-equivalence}
\begin{frame}{Wilf-equivalence}
  \begin{definition}
  Two patterns are said to be \emph{Wilf-equivalent} if the set of avoiders
  of the patterns is the same size at every length.
\end{definition}
 All coincident pattern classes are Wilf-equivalent.

 The permutations in \(\av{\perm{2,3,1}}\) of length \(n\) are counted by the
 Catalan numbers.
 \begin{equation}
   C_0 = 1, \text{ and } C_{n+1} =\sum_{i=0}^{n}C_iC_{n-i}
 \end{equation}
 This gives the sequence
 \begin{equation*}
   1,1,2,5,14,42,132,1430,4862,\dotsc
 \end{equation*}
\end{frame}
\subsection{Set Partitions}
\label{sub:Set Partitions}
\begin{frame}{Set Partitions}
  \begin{block}{Note}
    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
    \(\nrange{n}\). (Claesson \cite{A})
\end{block}
\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
\end{example}
All permutations in \(\av{\perm{2,3,1}}\) are also in \(\av{q}\).
\end{frame}
\begin{frame}
  Set Partitions can be used to explain certain Wilf-equivalences.
  \begin{example}
    The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
  \begin{equation*}
      m_1 = \pattern{}{1,2}{0/0,0/2,
                      1/0,1/1,
                      2/0,2/1,2/2}
      \text{ and }
      m_2 = \pattern{}{1,2}{0/0,
                      1/0,1/1,1/2,
                      2/0,2/1,2/2}
  \end{equation*}
  Consider containers of the patterns in \(\av{\perm{2,3,1}}\).

@@

\begin{frame}
  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\mathcal{M}_2}
\end{equation*}
Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\scriptstyle \mathcal{M}_2\setminus\varepsilon}
\end{equation*}
\end{frame}
\begin{frame}
  Both of these sets therefore have generating function \(M(x)\) satisfying
  \begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}\pause
 Solving \ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
\ref{eqn:Pgen} gives us that the generating function for
\(M(x)\) satisfies
\begin{equation}
    M(x) = xM^2(x) - x(M(x) - 1) + 1
\end{equation}\pause
Evaluating coefficients of \(M(x)\) then gives the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
\end{equation*}
This is an offset of the Motzkin numbers.
\end{frame}
